Supporting Information:\n prolfqua - A Comprehensive R-package for Proteomics Differential Expression Analysis

### 2022-10-22 inst/MyArticle/paper.Rmd
updated supplement.

By plotting the $TPR$ versus the $FPR$ we obtain the receiver operator characteristic curve (ROC curve) [@pROC2011]. The area under the curve ($AUC$) or partial areas under the curve ($pAUC$), at various values of the $FPR$, are measures of performance derived from the ROC curve. Using these measures, we can compare the performances of the statistics or the various models examined and test if these differences are statistically significant.

@@

knitr::asis_output(stringr::str_replace(tmp, "HACKIT", "$^{a}$"))

@@

```{r sessioninfo, echo=FALSE}
sessionInfo()
```

>>>>

By plotting the $TPR$ versus the $FPR$ we obtain the receiver operator characteristic curve (ROC curve) [@pROC2011]. The area under the curve ($AUC$) or partial areas under the curve ($pAUC$), at various values of the $FPR$, are measures of performance derived from the ROC curve. Using these measures, we can compare the performances of the statistics for a single model or the various models and test if the differences are statistically significant, using a test to compare ROC curves.

@@

knitr::asis_output(stringr::str_replace(tmp, "HACKIT", "$^{a}$"))
saveRDS(allBenchmarks, "allBenchmarks.RDS")

@@

```{r sessioninfo, echo=FALSE}
pander::pander(sessionInfo())
```

### 2022-10-22 inst/MyArticle/prolfqua_supplement.Rmd
updated supplement.

```{r installation, echo=TRUE, eval=FALSE}
remotes::install_github("https://github.com/fgcz/prolfqua/releases/tag/v0.4.3",
    dependencies = TRUE)
```

@@

The Rmarkdown file with R code to execute the benchmarks is  [Benchmarking of msqrob2, prolfqua and proDA using the CPTAC dataset MaxQuant output](https://wolski.github.io/prolfquabenchmark/articles/Benchmark_cptac.html)

@@

To build the vignettes with the benchmarks execute the following commands in the R console:

```{r installbenchmark, echo=TRUE, eval=FALSE}
install.packages('remotes')
remotes::install_gitlab("wolski/prolfquadata", host="gitlab.bfabric.org") # installs the data
remotes::install_github("wolski/prolfquabenchmark", build_vignettes = TRUE, dependencies = TRUE)
```

@@

```{r sessionInfo, echo=FALSE}
sessionInfo()
```

>>>>

```{r installation, echo=TRUE, eval=FALSE}
remotes::install_github("https://github.com/fgcz/prolfqua/releases/tag/v0.4.5",
    dependencies = TRUE, build_vignettes=TRUE)
```

```{r installation2, echo=TRUE, eval=FALSE}
remotes::install_gitlab("wolski/prolfquadata", host="gitlab.bfabric.org")
remotes::install_github("https://github.com/wolski/prolfquabenchmark",
    dependencies = TRUE, build_vignettes=TRUE)
```

# Comparing partial areas under the curve at 0.1 FDR for Ionstar/MaxQuant dataset

Here we test if we should reject the null hypothesis that the area under the ROC curve at 0.1 FDR does not differ for 
msqrob2, proDA, and prolfqua when computed based on scaled p.values. To this task we use the function roc.test from the R package `r BiocStyle::CRANpkg("pROC")`.

```{r roctest, include = FALSE, eval = TRUE}
allBenchmarks <- readRDS("allBenchmarks.RDS")
ttmsqrob <- allBenchmarks$benchmark_msqrob$data()
ddmsqrob <- pROC::roc(ttmsqrob$TP, ttmsqrob$scaled.p.value, partial.auc = c(1, 0.9))

ttprolfqua <- allBenchmarks$benchmark_merged$data()
ddprolfqua <- pROC::roc(ttprolfqua$TP, ttprolfqua$scaled.p.value, partial.auc = c(1, 0.9))
ttproda <- allBenchmarks$benchmark_proDA$data()
ddproda <- pROC::roc(ttproda$TP, ttproda$scaled.p.value, partial.auc = c(1, 0.9))

tmp <- c(msqrob2_vs_prolfqua = pROC::roc.test(ddmsqrob,ddprolfqua, progress = "none")$p.value,
         msqrob2_vs_proda = pROC::roc.test(ddmsqrob,ddproda, progress = "none")$p.value,
         prolfqua_vs_proda = pROC::roc.test(ddprolfqua,ddproda, progress = "none")$p.value)

The Table \@ref(tab:shwoROCtestREsultsIonStarMQ) shows the p-values of the bootstrap test for two ROC curves. We observe that for this benchmark data, there are no significant differences among the $AUC_{10}$ for the three packages.

```{r shwoROCtestREsultsIonStarMQ}
cap = "p-values for pairwise comparsions of partial AUC at 0.1 for msqrob2, proDA and prolfqua for IonStar MaxQuant data."
knitr::kable(tibble(pROC_test = names(tmp), p.value = round(tmp, digits = 2)), format = "latex", caption = cap)

@@

The Rmarkdown file which created these results, with more details about the CPTAC / MaxQuant dataset, can be found here:
["Benchmarking of msqrob2, prolfqua and proDA using the CPTAC dataset MaxQuant output"](https://wolski.github.io/prolfquabenchmark/articles/Benchmark_cptac.html)

```{r}
getpath <- function(filN){
    f1 <- paste0("../../inst/Benchresults/",filN)
    if (f1 == "") {
        f1 <- system.file("Benchresults",filN, package = "prolfquabenchmark")
    }
    message(f1)
    return(f1)
}
CPTACBenchmarks <- readRDS(getpath("CPTAC_Benchmark.RDS"))
mrob <- CPTACBenchmarks$benchmark_msqrob$pAUC_summaries()$ftable$content
mrob$package <- "msqrob2"
proda <- CPTACBenchmarks$benchmark_proDA$pAUC_summaries()$ftable$content
proda$package <- "proDA"
prolfqua <- CPTACBenchmarks$benchmark_merged$pAUC_summaries()$ftable$content
prolfqua$package <- "prolfqua"

all <- dplyr::bind_rows(list(mrob, proda, prolfqua))
all <- all |> filter(contrast == "b_vs_a")
all$what[all$what == "statistic"] <- "t_statistic"
all$what[all$what == "t"] <- "t_statistic"
all$what[all$what == "scaled.pval"] <- "scaled.p.value"
all$what[all$what == "logFC"] <- "diff"

Figure \@ref(fig:plotCPTACbarplot) shows the partial area under the curve for `r BiocStyle::Biocpkg("msqrob2")`, `r BiocStyle::Biocpkg("proDA")` and `r BiocStyle::Githubpkg("fgcz/prolfqua")`.

```{r plotCPTACbarplot, fig.cap= "Barplot showing the partial area under the ROC at 0.1 FDR."}
all |> ggplot(aes(x = what, y = pAUC_10, group = package, color = NULL, fill = package)) +
  geom_bar(stat = "identity",  position = position_dodge())

```{r roctestCPTAC, include = TRUE, message=FALSE, error=FALSE}

ttmsqrob <- CPTACBenchmarks$benchmark_msqrob$data()
ddmsqrob <- pROC::roc(ttmsqrob$TP, ttmsqrob$scaled.pval, partial.auc = c(1, 0.9))
ttprolfqua <- CPTACBenchmarks$benchmark_merged$data()
ddprolfqua <- pROC::roc(ttprolfqua$TP, ttprolfqua$scaled.p.value, partial.auc = c(1, 0.9))
ttproda <- CPTACBenchmarks$benchmark_proDA$data()
ddproda <- pROC::roc(ttproda$TP, ttproda$scaled.pval, partial.auc = c(1, 0.9))

tmp <- c(msqrob2_vs_prolfqua = pROC::roc.test(ddmsqrob,ddprolfqua, progress = "none")$p.value,
         msqrob2_vs_proda = pROC::roc.test(ddmsqrob,ddproda, progress = "none")$p.value,
         prolfqua_vs_proda = pROC::roc.test(ddprolfqua,ddproda, progress = "none")$p.value)

Table \@ref(tab:shwoROCtestREsultsCPTAC) shows the results of the Bootstrap test for two ROC curves. 
We observe that for this benchmark data, there are no significant differences among the $AUC_{10}$ for the three packages.

```{r shwoROCtestREsultsCPTAC}
cap = "p-values for pairwise comparsions of partial AUC for msqrob2, proDA and prolfqua CPTAC/MaxQuant data."
knitr::kable(tibble(pROC_test = names(tmp), p.value = round(tmp, digits = 2)), format = "latex", caption = cap)

# Benchmark IonQuant FragPipe v15 combined_protein.tsv

The Rmarkdown file which created these results, with more details about the IonQuant FragPipe v15 dataset, can be found here:
["Benchmarking proDA and prolfqua using Ionstar Dataset and FragPipe v15 compbined_protein output"](https://wolski.github.io/prolfquabenchmark/articles/Benchmark_cptac.html)

Direct comparison with `msqrob2` is impossible since, to fit the dropout model, the peptide intensities are required, while here, we are starting the analysis from the `combined_proteins.tsv` file.

```{r FragPipe}
FragPipeBenchmarks <- readRDS(getpath("FragPipev15_comb_prot.RDS"))
proda <- FragPipeBenchmarks$benchmark_proDA$pAUC_summaries()$ftable$content
proda$package <- "proDA"
prolfqua <- FragPipeBenchmarks$benchmark_prolfqua$pAUC_summaries()$ftable$content
prolfqua$package <- "prolfqua"

tmp <- dplyr::bind_rows(proda, prolfqua)
tmp$what |> unique()
tmp$what[tmp$what == "statistic"] <- "t_statistic"
tmp$what[tmp$what == "scaled.pval"] <- "scaled.p.value"
tmp <- tmp |> filter(contrast == "all")
```

Figure \@ref(fig:plotCPTACbarplot) shows the partial area under the curve for `r BiocStyle::Biocpkg("msqrob2")`, `r BiocStyle::Biocpkg("proDA")` and `r BiocStyle::Githubpkg("fgcz/prolfqua")`.

```{r plotFragPipebarplot, fig.cap= "Barplot showing the partial area under the ROC at 0.1 FDR."}
tmp |> ggplot2::ggplot(ggplot2::aes(x = what, y = pAUC_10, group = package, color = NULL, fill = package)) +
  ggplot2::geom_bar(stat = "identity",  position = ggplot2::position_dodge()) +
  ggplot2::facet_wrap(~ contrast) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, vjust = 0.5, hjust = 1))

```{r roctestFRAGPIPE, include = TRUE, message=FALSE, error=FALSE}

ttprolfqua <- FragPipeBenchmarks$benchmark_prolfqua$data()
ddprolfqua <- pROC::roc(ttprolfqua$TP, ttprolfqua$scaled.p.value, partial.auc = c(1, 0.9))
ttproda <- FragPipeBenchmarks$benchmark_proDA$data()
ddproda <- pROC::roc(ttproda$TP, ttproda$scaled.pval, partial.auc = c(1, 0.9))

tmp <- c(prolfqua_vs_proda = pROC::roc.test(ddprolfqua,ddproda, progress = "none")$p.value)

Table \@ref(tab:shwoROCtestREsultsFRAGPIPE) shows the results of the Bootstrap test for the difference of two ROC curves. We observe that for this benchmark data, there are no significant differences among the $AUC_{10}$ for the two packages.

```{r shwoROCtestREsultsFRAGPIPE}
cap = "p-values for pairwise comparsions of partial AUC for msqrob2, proDA and prolfqua CPTAC/MaxQuant data."
knitr::kable(tibble(pROC_test = names(tmp), p.value = round(tmp, digits = 2)), format = "latex", caption = cap)

@@

@@

```{r sessionInfo, echo=FALSE}
pander::pander(sessionInfo())
```
